Disclosing funding sources for open access publication fees: the Open APC

### 2016-05-26 article.Rmd
why we not enrich the other data-sets with crossref

We used cost data released by the Open APC initiative on May 13, 2016, to analyse expenditures on publication fees by German universities and research organisations. In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI.

Data collection in this study involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed  the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc. For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles. In the case of FWF, we gathered the cost information from accompanying reports. We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank.

Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOI's. Although the Open APC initative gathered metadata representing publishers and journals from CrossRef as well, this information was  retrieved at the time when the participating institutions submitted the cost information. 
However, CrossRef regularly updates metadata to represent ongoing mergers of publishing houses or name changes. A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that now operates as Springer Nature. To reflect these changes in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.

As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API. We requested the XML-based format `application/vnd.crossref.unixsd+xml`, which distinguished full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals. It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014]. In cases where no bibliographic information could be obtained, we used the Open APC values. Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.

>>>>

We analysed self-reported cost data released by the Open APC initiative on May 13, 2016, to assess institutional spending on open access publication fees in Germany. In addition to administrative data about the amount paid per article including value added tax, the reporting institution, and the year of invoicing, we used information about whether an article was published in a fully or hybrid open access journal as well as the recorded DOI from the data-set.

Instead of re-using publisher and journal information contained in the Open APC data-set, we fetched bibliographic metadata for each article from CrossRef on May 19, 2016, on the basis of the reported DOI's. Although the Open APC initiative gathered metadata representing publishers and journals from CrossRef as well, this information was  retrieved at the time when the participating institutions submitted the cost information. 
However, CrossRef regularly updates metadata to represent ongoing mergers of publishing houses or name changes. A prominent example in this regard was the merger of the two large publishing houses Springer Business + Media and Nature Publishing Group announced on May 6, 2015, that now operates as Springer Nature. To reflect these changes in academic publishing, we decided to retrieve updated metadata from CrossRef for the whole Open APC data-set.

As a client, we used the R package rcrossref [@rcrossref], developed and maintained by the rOpenSci initiative[^6], to access CrossRef's REST API. We requested the XML-based format `application/vnd.crossref.unixsd+xml` in which full and abbreviated journal titles as well as the media types of ISSN's, the International Standard Serial Number used to identify journals, were distinguished. It also contained disambiguated publisher information, thus avoiding confusion about naming of publisher houses other studies were faced with when working with self-reported data [@woodward_2014]. In cases where no bibliographic information could be obtained, we used the Open APC values. Since CrossRef is not the only registration agency for DOI's, but also the agencies DataCite and Medra mint DOIs for scholarly work, we furthermore obtained the DOI agency for each article with the help of the rcrossref client.

Data collection in this study also involved obtaining cost data from related open data initiatives. To compare self-reported spending on open access publishing by Germany universities and research organisations with that of other initiatives, we reviewed the openly available spreadsheets from the Wellcome Trust [@wellcome_apc], the Austrian Science Fund FWF [@fwf_apc] and Jisc. For analysis, we obtained the amount of overall spending on publications fees, as well as on charges to publish in fully open access journal articles. In the case of FWF, we gathered the cost information from accompanying reports. We used the spreadsheet data to summarize  Wellcome Trust's and Jisc's spending, and converted the prices from GBP to Euro in accordance with the average Euro foreign exchange reference rates provided by the European Central Bank. Our comparison between the open data initiatives focussed on the last two years 2014 and 2015.Since Wellcome Trust's spending was reported for the periods 2013 - 2014 and 2014 - 2015, we referred to the exchange rates of the full two-years as we could not determine the actual dates of spending. We excluded Jisc's spending in 2014 because full currency information in the data-set was not available.

Data collection methods of the Open APC initiative and those of the others differed in some aspects. For instance, whereas the DOI was a mandatory element in the Open APC data template that the participating institution were required to report, publication identifier in the Wellcome Trust data might also be added through automated compliance checks. For this reason, and as our main focus is institutional funding for publication fees in Germany, we decided only to compare German spending with that of other initiatives, but not its size and distribution over publisher and journal titles, as well as the indexing coverage in CrossRef. However, we made use of the literature to discuss our findings in relation to the other initiatives.
